---
title: "PostQC_OMIP095"
author: "Mikias HW"
date: "`r Sys.Date()`"
output: html_document
---
# Prep Workspace
## Load Libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(Spectre)
library(tidymodels)
tidymodels_prefer()
library(recipes)
library(flowCore)
library(FlowSOM)
library(janitor)
library(flowVS)
library(xlsx)
library(flowViz)
library(corrplot)
library(embed)
library(tidytext)
library(flowWorkspace)
library(CytoML)
library(ggcyto)
library(tidyverse)
library(flowAI)
library(bestNormalize)
library(patchwork)
library(ggforce)
library(SingleCellExperiment)
```

## Preferred functions
```{r results='hide', message=FALSE, warning=FALSE}
library(conflicted)
# Set function preferences (if needed)
conflicts_prefer(dplyr::select)

conflicts_prefer(dplyr::filter)
```

## Directories
```{r}
# Save working directory (if needed)
wDir <- getwd()

# Set working directory (if needed)
setwd(wDir)

# Set QC directory
qcDir <- paste(wDir, "FlowJo_QC/QC_Gating", sep = "/")

qcFCS <- paste(qcDir, "QC_FCS", sep = "/")

qcGates <- paste(wDir, "FlowJo_QC/FlowJo_Web_files", sep = "/")

# Save QC png list
qcPNG <- list.files(qcGates, pattern = "*.png", full=TRUE)

# flowAI results
resAI <- paste(wDir, "resultsQC", sep = "/")

```

# Project Background
## OMIP-095

    OMIP-095: 
    40-Color spectral flow cytometry delineates all major leukocyte populations in murine lymphoid tissues 

    Aris J. Kare, Lisa Nichols, Ricardo Zermeno, Marina N. Raie, Spencer K. Tumbale, Katherine W. Ferrara 

    First published: 28 September 2023 

    https://doi.org/10.1002/cyto.a.24788 
    
    Flow Repository ID: 	FR-FCM-Z63E
    
## Quality Control
```{r, results='asis'}
# Define markdown PNG display function
generate_PNGs <- function(png_paths) {
  image_html <- lapply(png_paths, function(path) {
    paste0("<img src=\"", path, "\" alt=\"Image\"/>\n")
  })
  return(unlist(image_html))
}

png_paths <- qcPNG

image_html <- generate_PNGs(png_paths)

cat(image_html, sep = "\n")

```

## Quality Control II
```{r eval=FALSE}
# # Select Flowjo Workspace file
# fjWSP <- list.files(qcDir, pattern="QC_Gated.wsp",full=TRUE)
# 
# # Select QC's FCS files
# FCSfiles <- list.files(qcFCS, pattern="*.fcs",full=TRUE)
# 
# # Create Flowjo Workspace object
# ws <- open_flowjo_xml(fjWSP)
# 
# # View sample information
# fj_ws_get_samples(ws, group_id = 1)
# 
# # Construct Gating Set object
# gs <- flowjo_to_gatingset(ws, group_id = 1)
# 
# # Exctact FCS parameters
# keys <- fj_ws_get_keywords(ws, 1)
# 
# # QC Steps 
# plot(gs)
# 
# gs_get_pop_paths(gs, path = "full")
# 
# gatelist <- gs_get_pop_paths(gs, path = "auto")
# 
# gate1 <- gatelist[1]
# gate2 <- gatelist[2]
# gate3 <- gatelist[3]
# gate4 <- gatelist[4]
# gate5 <- gatelist[5]
# 
# autoplot(gs, gate2, bins = 100)
# 
# gh <- gs[[5]]
# 
# autoplot(gh)
# 
# # Gating states
# gs_pop_get_stats(gs, xml = TRUE)
# 
# # Detectors
# gs_pop_get_data(gs)
# 
# # File names
# sampleNames(gs)

```

## Load FCS files
```{r}
# Create flowSet object
fs <- read.flowSet(
  path = qcFCS,
  pattern = "*.fcs",
  transformation = FALSE,
  truncate_max_range = FALSE
)

# rm(fs)
fs[[1]]

# get more info
pData(fs)
```

## flowAI final QC
```{r}
# # Create flowSet object
# fsPre <- read.flowSet(
#   path = qcFCS,
#   pattern = "*.fcs",
#   transformation = FALSE,
#   truncate_max_range = FALSE
# )
# 
# # rm(fs)
# fsPre[[1]]
# 
# # Can some time
# # Saves folder of QC results and QC'd FCS files in working directory
# 
# fs <- flow_auto_qc(fsPre, sideFM = "lower")

# Show QC results
QCfile <- paste(resAI, "QCmini.txt", sep = "/")

resAIqc <- read.csv(QCfile, sep = "\t", check.names = FALSE)

resAIqc

```


## Samples metadata
```{r}
# Prep empty df
metadata <- data.frame()

# Iterate through flowset to extract metadata
for (i in 1:length(fs)) {
  flow_frame <- fs[[i]]
  tube_name <- flow_frame@description[["TUBENAME"]] # Tube Name
  cell_count <- flow_frame@description[["$TOT"]] # Event Counts
  sample_group <- flow_frame@description[["GROUPNAME"]] # Sample Group
  fcs_version <- flow_frame@description[["FCSversion"]] # FCS File Version
  collection <- flow_frame@description[["$DATE"]] # Collection Date
  cytometer <- flow_frame@description[["$CYT"]] # Cytometer
  software <- flow_frame@description[["CREATOR"]] # Cytometer Software
  institute <- flow_frame@description[["$INST"]] # Institution
  operator <- flow_frame@description[["$OP"]] # Operator


  row_data <- data.frame(tube_name, cell_count, sample_group, 
                         fcs_version, collection, cytometer, 
                         software, institute, operator)
  metadata <- bind_rows(metadata, row_data)

  rm(flow_frame, row_data)
}

metadata

```

## Markers & Detectors lists
```{r}
# List detectors
pmts <- fs[[1]]@parameters@data[["name"]] 

# List markers
markers <- fs[[1]]@parameters@data[["desc"]] 

# Remove Time components
markers <- na.omit(markers)

pmts <- pmts[pmts != "Time"]

# Combine into dataframe
detList <- data.frame(markers, pmts) # Merge

detList
```

## Expression Data
```{r}
#fs[[1]]@exprs

exprsDat <- data.frame()

for (i in 1:length(fs)) {
  flow_frame <- fs[[i]]
  tube_name <- flow_frame@description[["TUBENAME"]] # Tube Name
  cell_exprs <- flow_frame@exprs # Expression Data

  row_data <- data.frame(tube_name, cell_exprs)
  exprsDat <- bind_rows(exprsDat, row_data)

  rm(flow_frame, row_data, cell_exprs, tube_name)
}

head(exprsDat)
```

## Subsample
```{r}

```

## Transformation
```{r}
# Apply Transformation paramets
BiExTrans <- biexponentialTransform("defaultBiexponentialTransform")

# Apply transformation to selected detectors
fsBiExTrans <- transform(fs, transformList(pmts, BiExTrans))

# ##############
ArcSinTrans <- arcsinhTransform("defaultArcsinhTransform", a=1, b=5)

# ##############
fsTransArc <- transform(fs, transformList(pmts, ArcSinTrans))

# Visualize CD25 before transformation
ggcyto(fs, aes(x = "CD25")) + 
  geom_freqpoly() + 
  labs(title = "Raw CD25")

ggcyto(fs, aes(x = "I-A_I-E")) + 
  geom_freqpoly() + 
  labs(title = "Raw I-A_I-E")

# Visualize CD25 after transformation.
ggcyto(fsBiExTrans, aes(x = "CD25")) + 
  geom_freqpoly() + 
  labs(title = "BioExponentially Transformed CD25")

ggcyto(fsBiExTrans, aes(x = "I-A_I-E")) + 
  geom_freqpoly() + 
  labs(title = "BioExponentially Transformed I-A/I-E")


```

## Visualize transformations
```{r}
# flowViz lattice
flowViz::densityplot(~`.`, fs[[1]]) 

flowViz::densityplot(~`.`, fsBiExTrans[[1]]) 

flowViz::densityplot(~`.`, fsTransArc[[1]])
```


## Extract transformed expression data
```{r}

exprsDatTrans <- data.frame()

for (i in 1:length(fsBiExTrans)) {
  flow_frame <- fsBiExTrans[[i]]
  tube_name <- flow_frame@description[["TUBENAME"]] # Tube Name
  cell_exprs <- flow_frame@exprs # Expression Data

  row_data <- data.frame(tube_name, cell_exprs)
  exprsDatTrans <- bind_rows(exprsDatTrans, row_data)

  rm(flow_frame, row_data, cell_exprs, tube_name)
}

head(exprsDatTrans)
```

## Modeling Clasifier
```{r}

# Split data into training/testing/validation sets
set.seed(1601)
fs_split <- initial_validation_split(exprsDatTrans, strata = tube_name, prop = c(0.025, 0.025))

fs_split

# Return data frames:
fs_train <- training(fs_split)
fs_test <- testing(fs_split)
fs_validation <- validation(fs_split)

# Check distribution of samples in train set
fs_train |>
  dplyr::count(tube_name)


set.seed(1602)

# Return an 'rset' object to use with the tune functions:
fs_val <- validation_set(fs_split)
fs_split$splits[[1]]

# Corrolation plot between all detectors
fs_train |>
  select(-tube_name) |>
  cor() |>
  corrplot(order = "hclust", type = "upper", diag = FALSE)

```


## Single Cell Experiment
```{r}
# Construct Single Cell Experiment Object
sce <- SingleCellExperiment(fs_train)

sce

# dim(assay(sce))
# colnames(colData(sce))
# colnames(rowData(sce))

```


## Tidy PCA
```{r}

# fs_rec <- recipe(tube_name ~ ., data = fs_train) |> 
#   step_zv(all_numeric_predictors()) |> 
#   step_orderNorm(all_numeric_predictors()) |>  
#   step_normalize(all_numeric_predictors())
# 
# fs_rec
# 
# fs_rec_trained <- prep(fs_rec)
# 
# fs_rec_trained
# 
# fs_rec_bakes <- bake(fs_rec_trained, fs_test)
# 
# 
# fs_rec_trained |> 
#   step_pca(all_numeric_predictors(), num_comp = 4) |> 
#   ggplot(aes(x = .panel_x, y = .panel_y)) +
#     geom_point(alpha = 0.4, size = 0.5)
# 
# 
# plot_validation_results <- function(recipe, dat = fs_validation) {
#   recipe %>%
#     # Estimate any additional steps
#     prep() %>%
#     # Process the data (the validation set by default)
#     bake(new_data = dat) %>%
#     # Create the scatterplot matrix
#     ggplot(aes(x = .panel_x, y = .panel_y, color = tube_name, fill = tube_name)) +
#     geom_point(alpha = 0.4, size = 0.5) +
#     geom_autodensity(alpha = .3) +
#     facet_matrix(vars(-tube_name), layer.diag = 2) + 
#     scale_color_brewer(palette = "Dark2") + 
#     scale_fill_brewer(palette = "Dark2")
# }
# 
# 
# 
# fs_rec_trained %>%
#   step_pca(all_numeric_predictors(), num_comp = 4) %>%
#   plot_validation_results() + 
#   ggtitle("Principal Component Analysis")

```

## flowSOM
```{r}

# # Run FlowSOM clustering
# resSOM <- FlowSOM(fsBiExTrans)
# 
# # Plot results
# PlotStars(resSOM,
#           backgroundValues = resSOM$metaclustering)
# 
# # Get metaclustering per cell
# flowSOM.clustering <- GetMetaclusters(resSOM)
# 


```


## Dimensionality reduction
```{r}
# 
# # Dimensionality reduction
# 
# cell.sub <- do.subsample(cell.dat, sub.targets, tissue.col)
# cell.sub
# 
# cell.sub <- run.umap(cell.sub, cluster.cols)
# cell.sub
# 
# # DR plots
# 
# make.colour.plot(cell.sub, "UMAP_X", "UMAP_Y", "FlowSOM_metacluster", col.type = 'factor', add.label = TRUE)
# make.multi.plot(cell.sub, "UMAP_X", "UMAP_Y", cellular.cols)
# make.multi.plot(cell.sub, "UMAP_X", "UMAP_Y", "FlowSOM_metacluster", tissue.col, col.type = 'factor')
# 
# # Expression heatmap
# 
# exp <- do.aggregate(cell.dat, cellular.cols, by = "FlowSOM_metacluster")
# make.pheatmap(exp, "FlowSOM_metacluster", cellular.cols)

```


